---
title: "Practical Machine Learning - Prediction Model for Weight Lifting Exercise Dataset"
author: "Gavin Kim"
date: "2016-12-15"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary
This is final report of Coursera Practical Machine Learning Course. Goal of this document is to build a prediction model of Weight Lifting Exercise. The dataset contains variables from sensor worn by users and the classes that type of exercise.
For match this goal, I used ensemble model of Generalized Boosted Regression Model, Random Forest, Conditional Inference Tree. And for reducing dimension to increase training speed, PCA is used. But the ensenble got no more performance of model than Random Forest. After all this approach, I could get 0.98% of accuracy.

## Loading data

#### Load libraries

```{r loadLib, message=FALSE, warning=FALSE}
library(caret)
library(randomForest)
library(corrplot)
library(MASS)

# For increasing learning speed, use multi-core.
library(doMC)
registerDoMC(cores = 8)
```

#### Download and Load training/test data

```{r loadData}
# Check existance of data directory and create it if not exist.
dataDir <- "data"
if(!dir.exists(dataDir))
  dir.create(dataDir)

# Download training and test file 
Url.train <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
Url.test <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
data.trainPath <- paste0(dataDir, "/", basename(Url.train))
data.testPath <- paste0(dataDir, "/", basename(Url.test))

if(!file.exists(data.trainPath))
  download.file(Url.train, data.trainPath, mode = "w")
if(!file.exists(data.testPath))
  download.file(Url.test, data.testPath, mode = "w")

data.train <- read.csv(data.trainPath)
data.test <- read.csv(data.testPath)
```

## Preprocessing data

After loaing data, we need to remove the columns what we don't use. There are columns that have only NA values and nearly zero variance. It's better that we could replace these columns from data and there's columns not related to predict weight lifting exercise like observation number or user name or timestamps. It also has to be removed.

```{r cleaningData}
# Remove columns having NA values
cs <- colSums(is.na(data.train))
data.train <- data.train[,(cs == 0)]
data.test <- data.test[,(cs == 0)]

# Remove near zero columns
nzvColumns <- nearZeroVar(data.train,saveMetrics=TRUE)
data.train <- data.train[,!nzvColumns$nzv]
data.test <- data.test[,!nzvColumns$nzv]

# Remove unnecessary columns (Username, timesstamps)
data.train <- data.train[,-c(1:6)]
data.test <- data.test[,-c(1:6)]
```

With checking correlation plot, we can get there's no strong relation between variables to disadventage building model.

```{r corr}
corrplot(cor(data.train[,-53]), method="pie", type="upper")
```

#### Split data to training and validation set

After cleaning data, I found the test data set has only 20 observasions. It's too small to check the accuracy of model.
Hence, I decided to split training data with training and validation. The validation data is for test model's accuracy and I'll use training data for building model with K-fold cross validation.

```{r splitData}
# split data for training/validation
inTrain <- createDataPartition(y=data.train$classe, p=0.7, list=FALSE)
data.trainPart <- data.train[inTrain,]
data.validPart <- data.train[-inTrain,]
```

## Modeling

I made plan to use ensemble model with stacking gbm, rf, ctree And PCA can reduce the size of variables for training performance.

```{r pca}
# 53th column is "classe" what will used for outcome label.
pcaModel <- preProcess(data.trainPart[,-53], method=c("pca", "center", "scale"), thresh = 0.95)
data.trainPca <- predict(pcaModel, data.trainPart[,-53])
data.trainPca <- data.frame(data.trainPca, classe = data.trainPart[,53])
data.validPca <- predict(pcaModel, data.validPart[,-53])
data.validPca <- data.frame(data.validPca, classe = data.validPart[,53])
print(pcaModel)
```

As you see above 25 components has captured with 95 percent of the variance.


```{r gbm}
set.seed(1234)
fit_gbm <- train(classe ~ ., data = data.trainPca, method = "gbm")
pred_gbm <- predict(fit_gbm, newdata = data.validPca)
confusionMatrix(pred_gbm,data.validPca$classe)
```

```{r rf}
set.seed(1234)
fit_rf <- train(classe ~ ., data = data.trainPca, method = "rf")
pred_rf <- predict(fit_rf, newdata = data.validPca)
confusionMatrix(pred_rf,data.validPca$classe)
```

```{r ctree}
set.seed(1234)
fit_ctree <- train(classe ~ ., data = data.trainPca, method = "ctree")
pred_ctree <- predict(fit_ctree, newdata = data.validPca)
confusionMatrix(pred_ctree,data.validPca$classe)
```

```{r ensemble}
predDF <- data.frame(pred_gbm, pred_rf, pred_ctree, classe = data.validPca$classe)
fit_stacked <- train(classe ~ ., data = predDF, method="rf")
pred_stacked <- predict(fit_stacked, newdata = predDF)
confusionMatrix(pred_stacked,data.validPca$classe)
```

The model what has best performance is Random forest as you see. And the ensemble model of gbm, rf, ctree gets no benefit of accuracy.


## Testing

We need answer to predict the 20 test set with this model.
Here is the results.

```{r testing}
# Apply PCA
data.testPCA <- predict(pcaModel, data.test[,-53])

# Create predicted variales
pred_gbm_test <- predict(fit_gbm, newdata = data.testPCA)
pred_rf_test <- predict(fit_rf, newdata = data.testPCA)
pred_ctree_test <- predict(fit_ctree, newdata = data.testPCA)

predDF_test <- data.frame(pred_gbm = pred_gbm_test, pred_rf = pred_rf_test, pred_ctree = pred_ctree_test)
pred_stacked_test <- predict(fit_stacked, newdata = predDF_test)
print(t(data.frame(pred_stacked_test)))

```

