---
title: "Practical Machine Learning - Prediction Model for Weight Lifting Exercise Dataset"
author: "Gavin Kim"
date: "2016-12-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading data

```{r loadData}
# Check existance of data directory and create it if not exist.
dataDir <- "data"
if(!dir.exists(dataDir))
  dir.create(dataDir)

# Download training and test file 
Url.train <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
Url.test <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
data.trainPath <- paste0(dataDir, "/", basename(Url.train))
data.testPath <- paste0(dataDir, "/", basename(Url.test))

if(!file.exists(data.trainPath))
  download.file(Url.train, data.trainPath, mode = "w")
if(!file.exists(data.testPath))
  download.file(Url.test, data.testPath, mode = "w")

data.train <- read.csv(data.trainPath)
data.test <- read.csv(data.testPath)
```

## Preprocessing data

```{r cleaningData}
library(caret)
# Remove columns having NA values
cs <- colSums(is.na(data.train))
data.train <- data.train[,(cs == 0)]
data.test <- data.test[,(cs == 0)]

# Remove near zero columns
nzv <- nearZeroVar(data.train,saveMetrics=TRUE)
data.train <- data.train[,!nzv$nzv]
data.test <- data.test[,!nzv$nzv]


str(data.train)
summary(data.train)
colSums(is.na(data.train))



```

After cleaning data, I found the test data set has only 20 observasions. It's too small to check the accuracy of model.
Hence, I decided to split training data with training and validation. The validation data is for test model's accuracy and I'll use training data for building model with K-fold cross validation.

```{r splitData}
# split data for training/validation
inTrain <- createDataPartition(y=training$classe, p=0.7, list=FALSE)
training_clean <- training_colRm[inTrain,]
validation_clean <- training_colRm[-inTrain,]

# PCA
res <- prcomp(data.train)
res$sdev
length(res$sdev)
res$rotation
dim(res$rotation)
res$x
dim(res$x)
preProcess( method="pca", pcaComp=)

head(data.train)

str(data.train)
summary(data.train)


train
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


## Check correlation between each features (Selecting Features)

## PCA


## Training
1. glm
2. gbm
3. rf
4. lda

assenble

## Validation

## Summary









